### 2020-04-05 Paper/Paper.tex
Updated paper and Main

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{ntheorem}
\newtheorem{hyp}{Hypothesis}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{
	Kho, Lee\\
	\textit{ltk224}
	\and
	Peeters, Jonas\\
	\textit{jp5642}
}
\title{DS-GA 1013: Project\\Intelligible Principal Component Analysis}

\usepackage[strings]{underscore}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{DS-GS 1013}
\lhead{Kho \& Peeters}
\chead{\thepage}

\setlength\parindent{0pt}

\input{macros}

\begin{document}
\maketitle
\section*{Deliverables}
\begin{enumerate}
\item Load the data
\item Performance benchmarking
\item PCA
\item Thresholding PCA
\item Simple Principal Components
\item CPEV versus Sparsity plots
\item SCoTLASS
\item SPCA
\end{enumerate}
\section*{Simple Principal Components}
\subsection*{Notes}
Vines (2000) generalized this approach without restricting the integer set solution using a sequence of simplicity preserving transformations. Each transformation chooses a pair of vectors and rotates them orthogonally in such a way that the variance associated with the currently higher variance component of the pair is increased at the expense of the lower variance component
\\\\
This method what introduced by Vines \citeyear{Vines2000}. We transform the set of orthogonal axes. Essentially we will be orthogonally rotating and rescaling two orthogonal vectors $d_1$ and $d_2$. In general we can write an orthogonal rotation as $f_1 = Pd_1$ where
\begin{equation}
P=\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}
\end{equation}
The new variance-covariance matrix $V^*$ of the rotated data can be found as $V^*=P^TVP$ where $V$ is the variance-covariance matrix of the data before rotation. 
\begin{align*}
V=\begin{bmatrix}
v_{11} & v_{12}\\v_{12} & v_{22}
\end{bmatrix}\quad \text{and}\quad V^* &= P^TVP=\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}^T\begin{bmatrix}
v_{11} & v_{12}\\v_{12} & v_{22}
\end{bmatrix}\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}\\
&= \begin{bmatrix}
h
\end{bmatrix}
\end{align*}
The variance in the direction $f_1$ can be found as follows:
\begin{equation}
\Var(Xf_1)=f_1^TVf_1=(Pd_1)^TVPd_1=d_1^TV^*d_1
\end{equation}

\pagebreak
\bibliographystyle{apacite}
\bibliography{references}
\end{document}