### 2020-04-15 Paper/Archive.tex
Updated Paper for Simple Components

\begin{equation}
P=\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}
\end{equation}
The new variance-covariance matrix $V^*$ of the rotated data can be found as $V^*=P^TVP$ where $V$ is the variance-covariance matrix of the data before rotation. 
\begin{align*}
V=\begin{bmatrix}
v_{11} & v_{12}\\v_{12} & v_{22}
\end{bmatrix}\quad \text{and}\quad V^* &= P^TVP=\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}^T\begin{bmatrix}
v_{11} & v_{12}\\v_{12} & v_{22}
\end{bmatrix}\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}
\end{align*}
The variance in the direction $f_1$ can be found as follows:
\begin{equation}
\Var(Xf_1)=f_1^TVf_1=(Pd_1)^TVPd_1=d_1^TV^*d_1
\end{equation}
essentially we want to find a $\beta^*$ such that the variance $\upnu$ in the direction of the new direction $f_1$ is maximized.
\begin{equation*}
\upnu=\frac{\|d_1\|_2^2\upnu_{11}+2\beta\|d_1\|\|d_2\|\upnu_{12}+\beta^2\|d_2\|^2\upnu_{22}}{\|d_1\|^2+\beta^2\|d_2\|^2}
\end{equation*}
for which we find that $\upnu$ is maximal if:
\begin{equation*}
\beta^*=\frac{-\|d_1\|\|d_2\|(\upnu_{11}-\upnu_{22})+\sqrt{\|d_1\|^2\|d_2\|^2(\upnu_{11}-\upnu_{22})^2+4\|d_1\|^2\|d_2\|^2\upnu_{12}^2}}{2\|d_2\|^2\upnu_{12}}
\end{equation*}