### 2020-04-15 Paper/Paper.tex

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{upgreek}
\usepackage{apacite}
\usepackage{ntheorem}
\newtheorem{hyp}{Hypothesis}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{
	Kho, Lee\\
	\textit{ltk224}
	\and
	Peeters, Jonas\\
	\textit{jp5642}
}
\title{DS-GA 1013: Project\\Intelligible Principal Component Analysis}

\usepackage[strings]{underscore}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{DS-GS 1013}
\lhead{Kho \& Peeters}
\chead{\thepage}

\setlength\parindent{0pt}

\input{macros}

\begin{document}
\maketitle
\section*{Deliverables}
\begin{enumerate}
\item Load the data
\item Performance benchmarking
\item PCA
\item Thresholding PCA
\item Simple Principal Components
\item CPEV versus Sparsity plots
\item SCoTLASS
\item SPCA
\end{enumerate}
\section*{Simple Principal Components}
Vines \citeyear{Vines2000} introduced the concept of simple principal component by restricting the magnitude of the loadings to integers such as $-1$, $0$ and $1$. This approached is based on the work for Hausman \citeyear{Hausman1982} who introduced a integer programming solution for assembling principal components using integers from a finite set. Each element is determined sequentially as to optimize the variance explained by each component. The problem with this method is that it rarely results in orthogonality over the components.\\
\\
Since it is hard to simplify principal components once they are calculated, Vines describes an alternative approach to obtaining simple components using an ortho-gonality preserving transformation on a pair of components. For each transformation of a pair of components, the variance of the data in the direction of one component is maximized while the variance in the other direction is minimized. This transformation is applied to a combination of principal components that has a high covariance. After this transformation, we exclude this pair of components for further transformations until all sensible transformations were made with the remaining components.\\
\\
During the "simplicity preserving transformation" introduced by Vines \citeyear{Vines2000}, we essentially orthogonally rotate and rescale two existing components $d_1$ and $d_2$. In general we can write an orthogonal rotation as $(f_1,f_2) = P(d_1,d_2)$ 
\begin{equation}
P=\begin{bmatrix}
1 & \|d_2\|_2^2 \beta\\\beta & -\|d_1\|_2^2
\end{bmatrix}
\end{equation}
V=\begin{bmatrix}
v_{11} & v_{12}\\v_{12} & v_{22}
\end{bmatrix}\quad \text{and}\quad V^* &= P^TVP

\\
\pagebreak
\bibliographystyle{apacite}
\bibliography{references}
\end{document}