### 2020-05-02 Presentation/Presentation.tex
Added beamer

\documentclass[11pt]{beamer}
\usetheme{Warsaw}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{
	Kho, Lee
	\and
	Peeters, Jonas
}
\title{Sparse Principal Components Analysis}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
%\logo{} 
\institute{\includegraphics{../Figures/nyu_long_black.png}} 
\date{\today} 
\subject{DS-GA 1013}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}\frametitle{Introduction}
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\begin{frame}\frametitle{State of the Art}
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Methodology}
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Results}
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Discussion}
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\end{document}